Post-replication citation patterns in psychology: Four case studies

### 2021-02-06 manuscript/manuscript (conflicted).Rmd
initial transfer of methods and results from google doc to r markdown complete

@@

```{r}
excluded_n <- d_contentAnalysis %>% filter(excluded == T) %>% nrow() 
excluded_reason <-  d_contentAnalysis %>% filter(excluded == T) %>% count(exclusionReason)
noAccess_n <- excluded_reason %>% filter(exclusionReason == 'No access') %>% pull(n)
nonEnglish_n <- excluded_reason %>% filter(exclusionReason == 'Non-English language') %>% pull(n)
refCiteOnly_n <- excluded_reason %>% filter(exclusionReason == 'Cited in references but not in text') %>% pull(n)
noCite_n <- excluded_reason %>% filter(exclusionReason == 'Does not cite target or replication') %>% pull(n)
```

We excluded `r excluded_n` citing articles from the qualitative analysis because we could not access the full text (n = `r noAccess_n`), they were non-English language (n = `r nonEnglish_n`), they included a citation to the original study in the reference section, but not in the main text (n = `r refCiteOnly_n`), or after manual inspection ßthey did not actually appear to cite the original study at all (n = `r noCite_n`).

@@

```{r}
artTyp <- d_contentAnalysis %>% 
  filter(excluded == F) %>% 
  count(articleType) %>%
  mutate(percent = round(n/sum(n),2)*100)

noData <- artTyp %>% filter(articleType == 'No empirical data')
metaAnalysis <- artTyp %>% filter(articleType == 'Data synthesis - meta-analysis')
caseStudy <- artTyp %>% filter(articleType == 'Empirical data - case study')
comment <- artTyp %>% filter(articleType == 'Empirical data - commentary including analysis')
fieldStudy <- artTyp %>% filter(articleType == 'Empirical data - field study')
labStudy <- artTyp %>% filter(articleType == 'Empirical data - laboratory study')
multipleStudy <- artTyp %>% filter(articleType == 'Empirical data - multiple study types are reported')
surveyStudy <- artTyp %>% filter(articleType == 'Empirical data - survey')
```

The research design of citing articles was classified as 'laboratory study' (n = `r labStudy$n`, `r labStudy$percent`%), 'no empirical data' (n = `r noData$n`, `r noData$percent`%), 'survey' (n = `r surveyStudy$n`, `r surveyStudy$percent`%), 'field study' (n = `r fieldStudy$n`, `r fieldStudy$percent`%), 'multiple study types' (n = `r multipleStudy$n`, `r multipleStudy$percent`%), 'meta-analysis' (n = `r metaAnalysis$n`, `r metaAnalysis$percent`%), 'commentary including data analysis' (n = `r comment$n`, `r comment$percent`%), or 'case study' (n = `r caseStudy$n`, `r caseStudy$percent`%).

@@

Figure \@ref(fig:citeCurves) shows standardized annual citation counts for each original study, the respective reference class (citations to all articles published in the same year and same journal as the original study), and classifications of citation valence (favourable, equivocal, unfavourable). The data can also be viewed in tabular format in (LINK TO APPENDIX). Valence classifications by the primary coder were modified after discussion with the secondary coder in `r classChange %>% filter(classificationChange == T) %>% pull(n)` (`r classChange %>% filter(classificationChange == T) %>% pull(percent)`%) cases.

>>>>

# Methods

The study protocol (rationale, methods, and analysis plan) was pre-registered on April 7th 2018 (https://osf.io/eh5qd/). An amended protocol was registered part way through data collection on May 1st 2019, primarily because we extended the sampling frame to cover additional months (https://osf.io/pdvb5/). All deviations from these protocols are explicitly acknowledged in Supplementary Information \@ref(appA). All data exclusions and measures conducted during this study are reported.

## Design
This was a retrospective observational study consisting of four case studies. Primary outcome variables were annual citation counts for original studies, citation valence (favourable, equivocal, unfavourable), co-citation of original and replication studies, frequency/type of counter-arguments.

## Sample
We examined four case studies in which a prominent pre-registered and multi-laboratory replication study strongly contradicted and outweighed the findings of an original study (Table 2).

```{r studyDetails, results = "asis"}
data_frame(
  'Original Study' = c('Baumeister et al. (May, 1998)','Sripada et al. (April, 2014)','Strack et al. (May, 1988)','Caruso et al. (July, 2012)','Carter et al. (July, 2011)'),
  'Replication study' = c('Hagger et al. (July, 2016)**','Hagger et al. (July, 2016)**','Wagenmakers et al. (October, 2016)','Klein et al. (January, 2014)','Klein et al. (January, 2014)'),
  'Effect' = c('Ego-depletion', 'Ego-depletion','Facial feedback','Money priming','Flag priming'),
  'Total citations to original*' = c('1974','36','708','57','54'),
  'Original sample size' = c('k = 1 N = 67', 'k = 1 N = 26', 'k = 1 N = 92', 'k = 1 N = 30','k = 1 N = 70'),
  'Replication sample size' = c('k = 23 N = 2,141', 'k = 23 N = 2,141', 'k = 17 N = 2,124', 'k = 36 N = 6,333','k = 36 N = 4,896'),
  'Original effect size [95\\% CI]' = c('d = 2.05 [1.31, 2.79]', 'd = 0.68 [0.09, 1.27]','MD = 0.82 [-0.05, 1.69]','d = 0.8 [0.05. 1.54]', 'd = 0.50 [.01, .99]'),
  'Replication effect size [95\\% CI]' =c('d = 0.04 [−0.07, 0.15]', 'd = 0.04 [−0.07, 0.15]', 'MD = 0.03 [-0.11, 0.16]', 'd = .01 [-.06, 0.09]', 'd = .01 [-.07, 0.08]')) %>% 
  kable(booktabs=T, longtable=T) %>% 
  kable_styling(latex_options="scale_down") %>%
  column_spec(1, width = "1em") %>%
  column_spec(2, width = "1em") %>%
  column_spec(3, width = "1em") %>%
  column_spec(4, width = "1em") %>%
  column_spec(5, width = "1em") %>%
  column_spec(6, width = "1em") %>%
  column_spec(7, width = "1em") %>%
  column_spec(8, width = "1em")
```

Valence classifications by the primary coder were modified after discussion with the secondary coder in `r classChange %>% filter(classificationChange == T) %>% pull(n)` (`r classChange %>% filter(classificationChange == T) %>% pull(percent)`%) cases.

@@

```{r}
excluded_n <- d_contentAnalysis %>% filter(excluded == T) %>% nrow() 
excluded_reason <-  d_contentAnalysis %>% filter(excluded == T) %>% count(exclusionReason)
noAccess_n <- excluded_reason %>% filter(exclusionReason == 'No access') %>% pull(n)
nonEnglish_n <- excluded_reason %>% filter(exclusionReason == 'Non-English language') %>% pull(n)
refCiteOnly_n <- excluded_reason %>% filter(exclusionReason == 'Cited in references but not in text') %>% pull(n)
noCite_n <- excluded_reason %>% filter(exclusionReason == 'Does not cite target or replication') %>% pull(n)
notexcluded_n <- d_contentAnalysis %>% filter(excluded == F) %>% nrow()
```

In total, `r d_citations %>% nrow()` articles cited one of the original studies of which `r d_contentAnalysis %>% nrow()` articles (after taking a 40% random sample in the Baumeister case) fell within the time period designated for qualitative assessment. Of these `r d_contentAnalysis %>% nrow()` articles, we excluded `r excluded_n` from the qualitative analysis because (1) we could not access the full text (n = `r noAccess_n`); (2) they were non-English language (n = `r nonEnglish_n`); (3) they included a citation to the original study in the reference section, but not in the main text (n = `r refCiteOnly_n`); or (4) manual inspection indicated that they did not actually appear to cite the original study at all (n = `r noCite_n`). Research design classifications for the remaining `r notexcluded_n` articles included in the qualitative analysis are shown in Table \@ref(tab:articleTypes).

@@

```{r articleTypes}
d_contentAnalysis %>% 
  filter(excluded == F) %>% 
  count(articleType) %>%
  mutate('Count (%)' = paste0(n,' (',round(n/sum(n),2)*100,')')) %>% 
  select('Article type' = articleType, 'Count (%)', -n) %>%
  apa_table(caption = 'Counts and percentages for research design classifications of articles included in qualitative analyses')
```

@@

Figure \@ref(fig:citeCurves) shows standardized annual citation counts for each original study and the respective reference class (citations to all articles published in the same year and same journal as the original study), and classifications of citation valence (favourable, equivocal, unfavourable, unclassifiable or excluded). The data can also be viewed in tabular format in Supplementary Table \@ref(tab:tabularData). All counts (n) reported in the text and table are raw counts (i.e., not standardized).
